Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/Response-MajRev/response_letter-version2.tex

\newenvironment{reply}
   {\medskip \noindent \textbf{Reply}:\  }
   {\medskip}

\newcommand{\shortreply}[2][]{\medskip \noindent \textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip}

\begin{document}

\section*{Response to the Reviewers}
% General intro text goes here
We thank the reviewers for their critical assessment and insightful comments on our work.
%
We have made extensive modifications to our manuscript.
%
In the following we address their concerns point by point.
%
%We also prepare an \emph{annotated version} for our revised manuscript.
%
%In the annotated version, changes corresponding to each point are all highlighted by red squares.

@@

TODO

% Let's start point-by-point with Reviewer 1
\reviewersection

% R1Q0
\begin{point}
    The authors conduct an empirical analysis of performance bottlenecks in graph neural network training. The authors identify the edge-related calculation is the performance bottleneck. Experimental of several GNN variants, such as GCN, GGNNNN, GAT and GaAN on six real-world graph datasets verify the importance of the findings. The experimental analysis is sufficient. However, there are some tiny issues in this paper.
\end{point}

\begin{reply}
    Thank you for your kindly comments on our manuscript.
    %
    We have revised the manuscript according to your suggestions. 
    %
    Please see the replies below.
\end{reply}

% Our reply
\begin{reply}
    We apologize for the confusing use of symbols.
    %
    To clarify the symbol usage, we have unified the usage of symbols in the revised manuscript.
    %
    After revision, each symbol only represents one meaning.

    Specifically, we use $\boldsymbol{s}$ to represent aggregated vectors in graph neurons.
    %
    We use $\boldsymbol{s}^l_x$ to denote the aggregated vector of the graph neuron corresponding to the vertex $v_x$ in the GNN layer $l$.
    %
    If the GNN layer $l$ has sub-layers, $\boldsymbol{s}^{l,i}_x$ represents the aggregated vector of $v_x$ in the $i$-th sub-layer.
    %
    As for edges and their feature vectors, we use $e_{y,x}$ to represent an edge pointing from $v_y$ to $v_x$ and we use $\boldsymbol{e}_{y,x}$ to represent its feature vector.

    In order to facilitate readers to quickly locate the meaning of  symbols, we have added Table~\ref{tab:symbols} at the beginning of Section 2 ``Review of Graph Neural Networks'' in the revised manuscript.

    \begin{table}[H]
        \footnotesize
        \caption{Frequently-used Symbols}
        \label{tab:symbols}
        \centering
        \begin{tabular}{p{3em}lp{35em}}
            \toprule
            Category & Symbol & Meaning \\
            \midrule
            \multirow[c]{4}{3em}{Graph Structure}& $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ & The simple undirected input graph with the vertex set $\mathcal{V}$ and the edge set $\mathcal{E}$. \\
            & $v_x$ & The $x$-th vertex of the input graph. \\
            & $e_{y,x}$ & The edge pointing from $v_y$ to $v_x$ of the input graph. \\
            & $\mathcal{N}(v_x)$ & The adjacency set of $v_x$ in the input graph. \\ 
            & $\bar{d}$ & The average degree of the input graph. \\ \midrule
            \multirow[c]{6}{3em}{GNN Definition}& $L$ & The number of GNN layers. \\
            & $K$ & The number of heads in a GNN layer. \\
            & $\phi^l$ & The messaging function of the GNN layer $l$. \\
            & $\Sigma^l$ & The aggregation function of the GNN layer $l$. \\
            & $\gamma^l$ & The vertex updating function of the GNN layer $l$. \\ 
            & $\phi^{l,i}$ / $\Sigma^{l,i}$ / $\gamma^{l,i}$ & The messaging/aggregation/updating function of the $i$-th sub-layer of the GNN layer  $l$.\\
            & $\textcolor{blue}{\boldsymbol{W}^l, \boldsymbol{W}^{(k)}/\boldsymbol{b}, \boldsymbol{a}}$ & The matrices/vectors represented by the blue characters are the weight matrices/vectors that need to be learned in the GNN. \\  \midrule
            \multirow[c]{8}{3em}{Vector}& $\boldsymbol{v}_x$ & The feature vector of the vertex $v_x$. \\
            & $\boldsymbol{e}_{y,x}$ & The feature vector of the edge $e_{y,x}$.  \\
            & $\boldsymbol{h}_x^{l}$ &  The {input} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$. \\
            & $\boldsymbol{h}_x^{l+1}$ &  The {output} hidden vector of the graph neuron corresponding to $v_x$ in the GNN layer $l$.\\
            & $\boldsymbol{m}_{y,x}^l$ & The message vector of the edge $e_{y,x}$ outputted by $\phi^l$ of the GNN layer $l$. \\
            & $\boldsymbol{s}_{x}^l$ & The aggregated vector of the vertex $v_x$ outputted by $\Sigma^l$ of the GNN layer $l$. \\
            & $\boldsymbol{h}_{x}^{l,i}$ / $\boldsymbol{m}_{y,x}^{l,i}$ / $\boldsymbol{s}_{x}^{l,i}$ & The hidden/message/aggregated vector of the vertex $v_x$ outputted by $\gamma^{l,i}$/$\phi^{l,i}$/$\Sigma^{l,i}$ of the $i$-th sub-layer of the GNN layer $l$. \\
            & $d^l_{in}$, $d^l_{out}$ &  The dimension of the input/output hidden vectors of the GNN layer $l$. \\
            & $dim(\MyVec{x})$ & The dimension of a vector $\MyVec{x}$. \\
            \bottomrule
        \end{tabular}
\end{table}

@@

\begin{reply}
	Thank you for pointing out missing references. Computer vision is indeed an important application area of graph neural networks. We have added the mentioned references in Section 1 ``Introduction'' in the revised manuscript.
\end{reply}

% R1Q3
\begin{point}
There are some grammar errors and typos:
\\
  - `Take the demo GNN in Figure 1(a) as  the example.'
\\
 - `to calculate the  output hidden vector $h^{l+1}$ of the current layer l, i.e., $h^{l+1}$ = $\gamma^l(h^l,s^l)$ The end-to-end training requires…'
\\
 - `Implementing it with the specially optimized basic operators on the GPU is a potential optimization'
\\
 - The sentences in the experimental section should be unified.
\end{point}

\begin{reply}

    Thank you for pointing them out.
    %
    We feel sorry for our carelessness.
    %
    We have proofread our revised manuscript to eliminate grammar errors and typos as much as we can.
    %
    We have also unified the tenses of the sentences in Section 3 ``Evaluation Design'' and Section 4 ``Evaluation Results and Analysis''.
    %
    Specifically, we use the past tense to describe experimental methods, results and what they indicate.
    %
    We use the present tense in the sentences that FigureX or Table X is the subject of the sentences.

@@

\end{reply}

% R1Q4
\begin{point}
    Figures 6 and 7 should be adjusted. The figures and fonts are too small.
\end{point}

\begin{reply}

    As suggested, we have split Figure 6 and Figure 7 into five figures (Figure 6 to 10) in the revised manuscript to enlarge their subfigures.
    %
    We have also adjusted font sizes in all figures (besides Figure 6 and Figure 7) to make sure that they are equal to or greater than the font size of figure captions.

@@

    \paragraph{Section 4.3 ``Memory Usage Analysis''}

     In the end of Section 4.3, we added a new paragraph ``Memory Usage of Inference'' to compare the memory usage of inference and training.
     %

     %
     Figure~\ref{fig:compare_memory_expasion_ratio} in the response (Figure X in the revised manuscript) compares the memory expansion ratios of typical GNNs during training and inference.
     %
     Since no intermediate results had to be cached during inference of GGNN, GAT, and GaAN, the memory expansion ratios of inference were much less than training for the three GNNs.
     %
     The MERs of inference were 45\% to 83\% (GGNN), 52\% to 61\% (GAT), and 37\% to 69\% (GaAN) of training.
     %
     However, the MERs were still high, disallowing inferencing with big graphs.
     %
     To handle big graphs, sample-based inference was necessary.

     \begin{figure}[H]
%        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{imagefile}}
     \end{figure}

@@

     \begin{figure}[H]
        \centering
        \subfloat[Batch Size = 1024]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_1024.pdf}}
        \subfloat[Batch Size = 2048]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_sampling_fix_batch_size_train_time_stack_2048.pdf}}
        \caption{Inference time per batch breakdown under different batch sizes.}
        \label{fig:time_breakdown_of_inference_sampler}
     \end{figure}
\end{reply}

% Begin a new reviewer section
\reviewersection

\begin{point}
    In the paper, authors accomplished a unique study and analysis on GNN models training complexity.  The articles first review and development history of GNNs and creatively model all architectures as input layers, intermediate layers of graph neurons and prediction layers. And they quantitatively summarize the time and space complexity of 4 representative GNNs, including graph convolution, gated recurrent graph net, graph attention net and GraphSage. Most importantly, the article first break down complexity into operator level and offered analysis of good granularity, giving reader more guidance in future study. At last, the solid experiments included the study of effects of hyper-parameters and a comparison of two major sampling techniques: neighbor sampling and cluster sampling.
\end{point}

\begin{reply}
    Thank you for your positive comments on our manuscript.
    %
    We have carefully revised the manuscript according to your kindly comments and suggestions in the following points.
    %
    %We highlight our modifications with red squares point by point in the annotated version of the revised manuscript.
\end{reply}

### 2020-12-25 NEUROC/sec4-experiments.tex
section4.3 update
更新了Effects on Accuracy

The relationship between hyper-parameters of GNNs (i.e., model complexity) and accuracy is very complicated, and there is no clear theoretical analysis yet.
%
Generally speaking, higher model complexity can bring powerful representation capability and may bring higher accuracy, but it also increases the risk of overfitting.
%
To evaluate the effects of hyper-parameters on accuracy, we measured the accuracy of the four typical GNNs under different hyper-parameters.
%
\figurename~\ref{fig:exp_hyperparameter_accuracy} shows the experimental results.

For GCN, the accuracy was sensitive to the dimension of hidden vectors $dim(\MyVec{h}^1_x)$.
%
As $dim(\MyVec{h}^1_x)$ increased, the accuracy first increased quickly and then stabilized when $dim(\MyVec{h}_x^1) \geq 8$.
%
For GGNN, its accuracy curves showed similar trends as GCN.
%
However, GGNN was more sensitive to $dim(\MyVec{h}_x^1)$ than GCN.
%
Its accuracy even decreased when $dim(\MyVec{h}_x^1) \geq 1024$.
%
Since GGNN had high model complexity (with 13 weight matrices/vectors to train), GGNN might occur overfitting in those cases.
%
For GAT, its accuracy was more sensitive to the dimension of each head $d_{head}$ than the number of heads.
%
For GaAN, only $dim(\MyVec{h}_x^1)$ showed obvious impacts on accuracy.
%
The experimental results indicated the accuracy of the GNNs was often low when $dim(\MyVec{h}_x^1)$ (for GCN/GGNN/GaAN) or $d_{head}$ (for GAT) was small.
%
In those cases, the low model complexity limited the expressive power of the GNNs.
%
As $dim(\MyVec{h}_x^1)$ or $d_{head}$ increased to a certain threshold, the GNNs had sufficient learning ability to achieve stable accuracy.

@@

%\begin{figure}[H]
    %\centering
    %\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat_small_info.pdf}
    %\caption{Effects of the number of heads on GAT ($d_{head}$=2,4)}
    %\label{fig:exp_hyperparameter_accuracy_gat_small_info}
%\end{figure}

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} compares the best accuracy that each GNN achieved on different datasets.
%
The best accuracy that the four GNNs achieved was very close.
%
It was also close to the accuracy reported in their original references \cite{shchur2018_pitfall_of_gnn, zeng2020_graphsaint}.
%
There was no clear winner in terms of accuracy.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN also achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
%
It indicated that simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

\subsubsection{Memory Usage of Inference}

@@

Since the size of sampled subgraph was much smaller than the original graph, it might affect the accuracy of the trained GNNs.
%
To find out how the batch size affected the accuracy, we trained GNNs under different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} compare their accuracy with the cases that we used the full graph participated in the training.
%
For each dataset and each GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section~\ref{sec:effects_of_hyper-parameters_on_performance}.

@@

The experimental results confirmed the effectiveness of the sampling methods in terms of model accuracy.
%
When the relative batch size was greater than or equal to 3\%, the accuracy of the model trained by sampling was close to the accuracy obtained by full-graph training.
%
In most cases, the accuracy achieved by the sampling methods was slightly lower than the full-graph training.
%
However, there were some exceptions (like GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc}), the accuracy achieved by sampling methods was even higher.

The relationships between batch size and accuracy were complex.
%
A large batch size did not always bring high accuracy.
%
For example, the accuracy of GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and GGNN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_fli} decreased as the batch size increased.
%
A small batch size sometimes could achieve higher accuracy.
%
For example, in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc}, the accuracy of GAT under 1\% relative batch size was even higher than the accuracy of the full-graph training.
%
Given a sampling method, we found that the optimal batch size highly depended on the dataset and the GNN algorithm.
%
Our observations were similar to \cite{zeng2020_graphsaint}.
%
How to automatically select the proper batch size is a topic worth further studying.

Among the two sampling methods, the performance of the cluster sampler was stabler than the neighbor sampler.
%
With the cluster sampler, the model accuracy of different GNNs was very close to the accuracy of full-graph training.
%
With the neighbor sampler, the model accuracy of different GNNs showed more obvious differences.

>>>>

%
The values of hyper-parameters determined the model complexity of a GNN.
%
The relationship between model complexity and accuracy was complex.
%
%As far as we know, there is little theoretical analysis yet.
%
Generally speaking, higher model complexity brought more powerful representation capability and might bring higher accuracy, but it also increased the risk of overfitting.

To evaluate the effects of hyper-parameters on accuracy, we measured the accuracy of the typical GNNs with varying hyper-parameters.
%
\figurename~\ref{fig:exp_hyperparameter_accuracy} (in the reply) shows the experimental results.
%
For GCN, its accuracy was sensitive to the dimension of hidden vectors $dim(\MyVec{h}^1_x)$.
%
As $dim(\MyVec{h}^1_x)$ increased, the accuracy first increased quickly and then stabilized when $dim(\MyVec{h}_x^1) \geq 8$.
%
For GGNN, its accuracy curves showed similar trends as GCN, but GGNN was more sensitive to $dim(\MyVec{h}_x^1)$ than GCN.
%
Its accuracy even decreased when $dim(\MyVec{h}_x^1) \geq 1024$.
%
Since GGNN had high model complexity (with 13 weight matrices/vectors to train), GGNN might occur overfitting in those cases.
%
For GAT, its accuracy was more sensitive to the dimension of each head $d_{head}$ than the number of heads.
%
For GaAN, only $dim(\MyVec{h}_x^1)$ showed obvious impacts on accuracy.
%
The experimental results indicated that the accuracy of the GNNs was often low when $dim(\MyVec{h}_x^1)$ (for GCN/GGNN/GaAN) or $d_{head}$ (for GAT) was too low.
%
%The low model complexity limited the representation capability of the GNNs.
%
As $dim(\MyVec{h}_x^1)$ or $d_{head}$ increased to a certain threshold, the GNNs gained sufficient learning ability to achieve stable accuracy.

@@

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} further compares the best accuracy that each GNN could achieve on different datasets.
%
The best accuracy of the typical GNNs was very close.
%
%
It was also close to the accuracy reported in their original references [34, 31].
%
The results indicated that there was no clear winner.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN achieved the highest or second highest accuracy in three out of five datasets, though its model was simplest.
%
Simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

\paragraph{Memory Usage of Inference}

@@

Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models got by sampling techniques might be different from the full-batch training.
%
To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} (in the reply) and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} (in the reply) compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
%
For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.

@@

The experimental results confirmed the effectiveness of the sampling methods in terms of accuracy.
%
When the relative batch size was greater than or equal to 3\%, the test accuracy of the GNNs trained with sampling was close to the accuracy obtained by full-batch training.
%
In most cases, the accuracy achieved by the sampling techniques was slightly lower than the full-batch training.
%
However, there were some exceptions (like GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc} in the reply) that the accuracy achieved by sampling was even higher.

The relationships between batch size and test accuracy were complex.
%
A larger batch size did not always bring higher accuracy.
%
For example, the accuracy of GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} (in the reply) and GGNN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_fli} (in the reply) decreased as the batch size increased.
%
A smaller batch size sometimes could achieve higher accuracy.
%
For example, GAT achieved a higher accuracy with 1\% relative batch size than the full-batch training in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc} (in the reply).
%
Given a sampling method, we found that the optimal batch size highly depended on the dataset and the GNN algorithm.
%
Our observations were similar to [31].
%
How to automatically select a proper batch size is a topic worth further studying.

Among the two sampling methods, the performance of the cluster sampler was stabler than the neighbor sampler.
%
With the cluster sampler, the test accuracy of different GNNs was very close to the accuracy of full-batch training.
%
With the neighbor sampler, the test accuracy of different GNNs showed more obvious differences.
